---
title: "ExploratoryAnalyses_warbleR"
output: html_notebook
---

Loading latest libraries for warbleR and Rraven using devtools
```{r}
devtools::install_github("maRce10/warbleR")
library(warbleR)
library(dplyr)
library(Rraven)

```

Importing selection tables and annotations made from Raven

Steps taken to create a selection table using Raven:
1. Load a given recording using Raven Pro 1.6.1.
2. Select a given feature and save it within your training data folder
3. Re-load the file - one at a time 
2. Create a selection table by dragging and dropping a selection box around the feature of interest.
3. Press enter to commit the selection table.
4. Right-click on the tool bar and select 'Choose Measurements'.
5. Select the following measurements to ensure that a selection table is loaded in R and recognized by the Rraven package: selection, view, channel, beginTime(s), endTime(s), lowFreq(Hz), highFreq(Hz), Begin Path, Begin File, File Offset(s), Beg File Samp (samples)
6. Click File - Save selection table as - save it in a folder that contains the acoustic data as well (.wav file)



```{r}

rvn.dat <- imp_raven(all.data=T, sound.file.col = "sound.files", warbler.format =  TRUE, path ="C:\\Users\\Vijay\\Desktop\\Chapter 3\\Data\\Summer 2020\\wcb\\")

rvn.dat.st <- selection_table(rvn.dat[1,], path ="C:\\Users\\Vijay\\Desktop\\Chapter 3\\Data\\Summer 2020\\wcb\\")


trial <- selection_table(basename(a), path="C:\\Users\\Vijay\\Desktop\\Chapter 3\\Data\\Summer 2020\\wcb\\" )

```
















Load recordings from xeno-canto and create a map
```{r}
wcb <- quer_xc("Psilopogon viridis",download = F, parallel = 4)
xcmaps(wcb, img = FALSE) 

```


Load the necessary training data and create a selection table
```{r}
# Create a vector of all the recordings in the directory
list.wav.file.names <- list.files("C:\\Users\\Vijay\\Desktop\\Chapter 3\\Data\\Summer 2020\\wcb\\", full.names=T)

# What's the duration of these recordings?
duration <- wavdur(files=basename(list.wav.file.names), path = "C:\\Users\\Vijay\\Desktop\\Chapter 3\\Data\\Summer 2020\\wcb\\")

# Create long spectrograms with all the training data
long_spec <- lspec(flist = basename(list.wav.file.names), ovlp = 10, it = "tiff",
                    path = "C:\\Users\\Vijay\\Desktop\\Chapter 3\\Data\\Summer 2020\\wcb\\")

# Now create selection tables
data <- duration %>%
  mutate(start= 0, end = duration, selec = 1)
data <- data[,-2]  
colnames(data) <- c("sound.files","start","end","selec")

dat_table <- selection_table(data,path = "C:\\Users\\Vijay\\Desktop\\Chapter 3\\Data\\Summer 2020\\wcb\\")

```


Carry out spectrographic cross correlation
```{r}
# Load the acoustic data to be matched with
a <- tuneR::readWave("C:\\Users\\Vijay\\Desktop\\Chapter 3\\HP36P3B\\20200321_090000.wav")
test <- file.path(tempdir(), "test.wav")
writeWave(a, test)

trial <- 


# make matrix with templates and sound files
cm <- matrix(c(dat_table[1],test), ncol = 2)

cm <- matrix(c("Phae.long1.wav-1", "Phae.long1.wav"), ncol = 2)
             
t <- cbind(dat_table[,1],rep(test,nrow(dat_table)))
matrix(c(dat_table[,1],test),ncol=2)


xc.output <- xcorr(dat_table, bp=c(0,2),
                   output = "list", path = "C:\\Users\\Vijay\\Desktop\\Chapter 3\\Data\\Summer 2020\\wcb\\")

pks <- find_peaks(xc.output = xc.output, path = "C:\\Users\\Vijay\\Desktop\\Chapter 3\\Data\\Summer 2020\\wcb\\")

```


# Create empty list
list.of.sound.files <-  list()

# Loop to read in all .wav files in directory and add the file names and .wav files to a list 
for (x in 1:length(list.wav.file.names)) { 
  tmp.wav <- warbleR::read_wave(basename(list.wav.file.names[x]),
                                path = "C:\\Users\\Vijay\\Desktop\\Chapter 3\\Data\\Summer 2020\\wcb\\")
  list.of.sound.files[[x]] <- list(list.wav.file.names[x], tmp.wav) 
}

# Create a selection table

list.of.sound.files[[1]]


trial <- tuneR::readWave(list.wav.file.names[1], header=TRUE)


data(list = c("Phae.long1", "Phae.long2", 
              "Phae.long3", "Phae.long4", "selec.table"))

#write files to disk
writeWave(Phae.long1,"Phae.long1.wav")

cm <- matrix(c("Phae.long1.wav-1", "Phae.long1.wav"), ncol = 2)

# Run cross correlation
xc.output <- xcorr(X = lbh_selec_table[1, ], output = "list", 
compare.matrix = cm, path = tempdir())

```

