---
title: "03_mean-spectra-freq-peaks"
output: html_notebook
---

Install necessary librarie
```{r}
library(seewave)
library(warbleR)
library(tuneR)
```


Exploratory data analysis - Here I will try a few different functions from the seewave package in R to get a better sense of the distribution of frequencies over time
```{r}
library(seewave)

# Read a sample .wav file
a <- tuneR::readWave("data\\20200307_083000.wav")

# Calculate mean frequency spectrum
# This function returns the mean frequency spectrum (i.e. the mean relative amplitude of the frequency distribution) of a time wave. Results can be expressed either in absolute or dB data.
mf <- meanspec(a,f=48000,wl=256,norm=F)

# Calculate frequency peaks
# This function searches for peaks of a frequency spectrum.
f_peak <- fpeaks(mf, threshold = 0.003)

# Frequency bands plot
# This graphical function returns a frequency spectrum as a bar plot.
f_bands <- fbands(mf)

# Soundscape frequency spectrum of a time wave
# This function returns a kHz binned spectrum as described by Kasten et al. (2012) for the description of a soundscape.
# This function essentially creates a bar plot of frequencies in kHZ by amplitude values
s_spectr <- soundscapespec(a)

```

Borrowing a function from the soundecology package in R and modifying it to suite our needs
```{r}
# Function that gets the proportion of values over a db value in a specific band of frequencies. Frequency is in Hz
getscore <- function(spectrum, minf, maxf, db, freq_row){
		miny<-round((minf)/freq_row)
		maxy<-round((maxf)/freq_row)
		
		subA = spectrum[miny:maxy,]
		
		index1 <- length(subA[subA>db]) / length(subA)
		
		return(index1)
	}

# Window length for the spectro and spec functions to keep each row every 10Hz
# Frequencies and seconds covered by each
freq_per_row <- 10
samplingrate <- 48000
wlen <-  samplingrate/freq_per_row # Window length

# This function returns a two-dimension spectrographic representation of a time wave. The function corresponds to short-term Fourier transform. An amplitude contour plot can be overlaid.
spec <- spectro(a, f = samplingrate, wl = wlen, plot = FALSE)$amp

# Set the max freqency that can be used
# This is often the nyquist frequency which is calculated as sampling frequency/2
max_freq <- samplingrate/2 

# Set an amplitude threshold
db_threshold <- -50

# Setting freqeuncy steps
freq_step <- 1000

# A sequence of frequencies
Freq <- seq(from = 0, to = max_freq - freq_step, by = freq_step)

Score <- rep(NA, length(Freq))
for (i in 1:length(Freq)) {
			Score[i] = getscore(spec, Freq[i], (Freq[i] + freq_step), db_threshold, freq_per_row)
		}


```




Campos-Cerqueira et al., 2019 inputs:
For meanspec: f=44,100, wl = 256, wn = ‘hanning’, norm = FALSE
Audio waveforms scaled between -1 and 1
Spectral peaks limited to max. amplitude of 1
Amplitude threshold = 0.003
Counted number of recordings with a peak in each of the 128 frequency bins (given amplitude threshold)