---
title: "04-ordination-NMDS-analysis"
output: html_document
---

This script is for site-wise (Benchmark, Actice and Passive sites) species composition ordination analysis for the 2020_summer_annotated data.


Install required packages
```{r}
library(tidyverse)
library(dplyr)
library(stringr)
library(vegan)
library(tidyr)
```

First step is to Sort the df from the annotated datasheet for running ordination (NMDS) analysis.

```{r}
#Attach the 2020_summer_annotation data 
data <- read.csv("data\\2020_Summer_annotation_AR_ver_01.csv")
```


#Site-wise (4 minute) sorting 
```{r}
#Split the Filenames into 4 columns : Site, Date, Time and Splits   
data1 <- separate(data, col = Filename, into = c("Site", "Date", "Time", "Splits"), sep = "_")

data2 <- data1 %>% group_by(Site, Date, Time, Restoration.Type..Benchmark.Active.Passive.) %>% summarise_at(.vars = vars(c("IP":"ISO")),
.funs = sum)
data2

```

In terms of next steps:
1. I would say - load in a combined data file that contains all your annotations as well as the latest one that includes Akshay's - let's call it as our working document or something
2. repeat the steps you did above.
3. For every column whose value is >1, count it as 1 - that's for species richness right? For example GJ in the first row has a value of 14, which means there were 14 annotations for the first 5 minutes, but species richness should be 1. 
4. Once that is done, you can do perhaps explore a few metrics:
- for example, group_by SITEID alone - gives you an overall species richness by SITE alone
- group by treatment type - gives you species richness by benchmark/restored/unrestored.

5. You can take a look at Priyanka's code to then calculate some other general metrics once the data has been formatted in a way it is needed. 


