---
editor_options: 
  chunk_output_type: console
---

# Splitting large .wav files

Here, we will first split the raw data which was collected for 24 hours at a site, for 7 days at a stretch. This is being done for the sake of manual annotation of bird species. The deployment schedule of the AudioMoths was set to record for 4-minutes and was switched off for 1-min. For the sake of analysis, data was split into 10s chunks and annotated manually using Raven Pro. 

### Load required libraries
```{r}
library(warbleR)
library(seewave)
library(dplyr)
library(stringr)
library(tools)

# Source any custom/other internal functions necessary for analysis
source("code/01_internal-functions.R")
```

### Selecting dawn acoustic data

We will use warbleR::split.wavs() to split a large file. To do so, we will first load a list of .wav files from folders (will have to be done site by site). Next, we we select only files between 6 am and 10 am (this can be varied depending on the exercise or the question at hand). For each day selected, we randomly extracted a continuous 16-min of recording.  
```{r}
# List the path that contains all folders, which contain the audiomoth data
path <- "C:\\data\\2020-winter\\"

# Listing the folders within which .WAV files are stored
folders <- dir(path, recursive=F,full.names=T)

# Now get only those files that begin at 6am and end at 10am
files <- list()

for(i in 1:length(folders)){

setwd(folders[i])

# Below code needs to be run only if we have to rename files
# List the files within each folder and renaming the files with the prefix - SITE_ID
a <- list.files(paste0(path,basename(folders)[i],"\\"), full.names = T)
file.rename(from = a, to=paste0(basename(folders)[i],"_",basename(a)))

# Extract the strings for .wav files between 6am and 10am
time_str <- list.files(paste0(path,basename(folders)[i],"\\"),full.names = T) %>%
                    tools::file_path_sans_ext() %>% str_extract('\\d+$')
time_str <- time_str[time_str>="060000" & time_str <="100000"] # vary times here depending on the question at hand

  for(j in 1:length(unique(time_str))){
    
    b <- list.files(paste0(path,basename(folders)[i],"\\"),full.names = T,
                    pattern = time_str[j])
    files <- c(files,b)
    }
}

# These are the list of files we need
files <- unlist(files)

# Now we choose a random consecutive 16 min of data between 6am and 10am
# Get a list of unique dates (since we will be generating a random 16min for every date across every site)

site_date <- str_extract(basename(files),'\\w+_\\d+_')
unique(site_date) # Give you unique date and sites for which we need to generate 16 min of data

subset_files <- list()

for(i in 1:length(unique(site_date))){
  a <- files[str_detect(files,unique(site_date)[i])]
  if(length(a)<4){ # essentially specifies that the min number you need
    next
  } else {
  subset_dat <- extractRandWindow(a,4)
  subset_dat <- na.exclude(subset_dat) # If there are less than 4 files
  subset_files <- c(subset_files, subset_dat)
  }
}

final_subset <- unlist(subset_files)

# Subset those files and copy it to a separate folder
# Please note that these folders & files are locally stored (they are extremely large and cannot be added to GitHub)
dir.create(paste0("C:\\data\\","subset"))
file.copy(from = final_subset, to="C:\\data\\subset\\")
```

### Split the files

Split the files and provide unique names to each file
```{r}
# Note: the path you choose to store data is upto the user.
subset_path <- "C:\\data\\subset\\"

# Split the files into n-second chunks
split_wavs(path=subset_path, sgmt.dur = 10, parallel=4)

# Get files that need to be renamed
split_files <- list.files(subset_path, full.names = T, pattern = "-")

# Note the number of chunks will vary as a function of segment duration
# 240 seconds = 24 chunks each of 10s

setwd(subset_path)

chunks <- c("01-10","10-20","20-30",
            "30-40","40-50","50-60",
            "60-70","70-80","80-90",
            "90-100","100-110","110-120",
            "120-130","130-140","140-150",
            "150-160","160-170","170-180",
            "180-190","190-200","200-210",
            "210-220","220-230","230-240")

for(i in 1:length(chunks)){
  c <- split_files[endsWith(split_files,paste0("-",i,".wav"))]
  d <- str_replace(c,paste0("-",i),paste0("_",chunks[i]))
  file.rename(from=c, to=d)
}

# Remove the original files
orig_files <- list.files(subset_path, full.names = T, pattern = ".WAV$")
file.remove(orig_files)
```

Now, go ahead and begin the process of manual annotation!