---
editor_options: 
  chunk_output_type: console
---

# Multiple regression of distance matrices


Load necessary libraries
```{r}
library(tidyverse)
library(dplyr)
library(stringr)
library(vegan)
library(ggplot2)
library(scico)
library(sf)
library(sna)
library(hier.part)

# Source any custom/other internal functions necessary for analysis
source("code/01_internal-functions.R")
```


Load necessary data and create dissimilarity matrices to run Mantel's tests
```{r}
# estimating geographic distance (euclidean) between sites
sites <- read.csv("data/list-of-sites.csv")

sites <- st_as_sf(sites, coords = c("Longitude","Latitude"), crs=4326)
sites <- st_transform(sites, 32643)

distanceData <- dist(st_coordinates(sites), method = "euclidean", diag = TRUE, upper = TRUE, p = 2)
distMatrix <- as.matrix(distanceData)

# getting vegetation PCA scores as a matrix
vegPcaScores <- read.csv("data/pcaVeg.csv")
vegPcaScores <- dist(vegPcaScores, method = "euclidean", diag = TRUE, upper = TRUE, p = 2)
vegPcaScores <- as.matrix(vegPcaScores)

# load NMDS scores (run on bird detections)
nmdsScores <- read.csv("data/nmdsBrayCurtis.csv")
birdMatrix <- dist(nmdsScores[,1:2], method = "euclidean", diag = TRUE, upper = TRUE, p = 2)
birdMatrix <- as.matrix(birdMatrix)
```


Extracting floristic data and running NMDS ordinations on the same
```{r}
# getting floristic data as a matrix, using NMDS scores
# to do the above, we need to reload the vegetation data and process it

veg <- read.csv("data/2020-vegetation-data.csv")
veg$Site_ID <- str_remove(veg$Site_ID,"_")

# We load the subset data
datSubset <- read.csv("data/datSubset.csv")
sites_needed <- data.frame(unique(datSubset$Site))
names(sites_needed) <- "Site_ID"

# Obtain a subset of the data which have the sites visited 
veg <- merge(veg, sites_needed, by.x = "Site_ID", by.y="Site_ID")

# renaming restoration type
veg$Site_type[veg$Site_type=="Unrestored"] <- "Passive"
veg$Site_type[veg$Site_type=="Restored"] <- "Active"

# now extract floristics data (tree species abundance)
floraData <- veg %>% 
  dplyr::select(Site_ID, tree_species, Site_type) %>% 
  mutate(Number = 1) %>%
  group_by(tree_species, Site_ID, Site_type) %>% 
  summarise (totalAbun = sum(Number)) %>% 
  pivot_wider (names_from = tree_species, values_from = totalAbun, values_fill = list(totalAbun=0)) %>%
  arrange(.,Site_ID)

# Convert to matrix form
nmdsDatMatrix <- as.matrix(floraData[, 3:ncol(floraData)])

# Run a bray-curtis dissimilarity index and use metaMDS function from vegan to run ordinations
disBrayCurtis <- vegdist(nmdsDatMatrix, method = "bray")

nmdsBrayCurtis <- vegdist (nmdsDatMatrix, method = "bray") %>% 
  metaMDS (nmdsBrayCurtis, k=6)

# extract nmds scores
nmdsScores <- as.tibble(scores(nmdsBrayCurtis))

# stress is lowest at around 6 dimensions with value of 0.06350522. However, if stress is high, we should reposition the points in 2 dimensions in the direction of decreasing stress, and repeat until stress is below some threshold.**A good rule of thumb: stress < 0.05 provides an excellent representation in reduced dimensions, < 0.1 is great, < 0.2 is good/ok, and stress < 0.3 provides a poor representation.** To reiterate: high stress is bad, low stress is good!

# create a distance matrix on the nmds scores of the flora data
floraMatrix <- dist(nmdsScores[,1:2], method = "euclidean", diag = TRUE, upper = TRUE, p = 2)
floraMatrix <- as.matrix(floraMatrix)
```


Running mantel's tests on the dissimilarity matrices of bird detections (nmds scores), vegetation (pca scores) and flora abundance (nmds scores)
```{r}
mantelBirdPca <- vegan::mantel(birdMatrix, vegPcaScores)
mantelBirdFlora <- vegan::mantel(birdMatrix, floraMatrix)

# Mantel's and partial mantel's tests suggest no strong relationship

vegan::mantel.partial(birdMatrix, vegPcaScores, distMatrix)
vegan::mantel.partial(birdMatrix, vegPcaScores, floraMatrix)
```


Running multiple regression on distance matrices

```{r}
# First we get rid of redundant duplicate values, then make the diagonal zeros NA, unfold each matrix into a vector, and then omit rows with 0 in them

birdVector <- upper.tri.remove(birdMatrix)
diag(birdVector) <- NA
birdVector <- cbind(c(birdVector)) %>% na.omit()

floraVector <- upper.tri.remove(floraMatrix)
diag(floraVector) <- NA
floraVector <- cbind(c(floraVector)) %>% na.omit()

pcaVector <- upper.tri.remove(vegPcaScores)
diag(pcaVector) <- NA
pcaVector <- cbind(c(pcaVector)) %>% na.omit() 

distanceVector <- upper.tri.remove(distMatrix)
diag(distanceVector) <- NA
distanceVector <- cbind(c(distanceVector)) %>% na.omit() %>%
  scale

#arranging data for MRM
mrmData <- bind_cols (birdVector, floraVector, pcaVector, distanceVector) %>%
  rename (bird = ...1, flora = ...2, structure = ...3, distance = ...4)

MRM(dist(bird) ~ dist(flora) + dist(structure) + dist(distance), data = mrmData, method = "linear", nperm = 1000)

# The above result suggests that bird compositional dissimilarity is not signficantly related to dissimilarity in habitat structure, geographic distance and florstics (Please note that the overall R2 values are very low)
```


Hierarchical partitioning

```{r}
hier.part(mrmData$bird, mrmData[2:4], family = "gaussian", gof = "Rsqu", barplot = TRUE)
```

