---
editor_options: 
  chunk_output_type: console
---

## Generalized linear modeling (acoustic space use and proportion of bird species detections and time since restoration)

In this script, we run generalized linear models to test the association between acoustic space use values and species richness, and time since restoration. 

### Install required libraries
```{r}
library(tidyverse)
library(dplyr)
library(stringr)
library(vegan)
library(ggplot2)
library(scico)
library(psych)
library(rcompanion)
library(multcomp)
library(lme4)
library(ggpubr)
library(sjPlot)

# Source any custom/other internal functions necessary for analysis
source("code/01_internal-functions.R")
```

### Load necessary data for statistical modeling
```{r}
# load list of sites
sites <- read.csv("data/list-of-sites.csv") %>% 
  dplyr::select("Site.code","Restoration.type") %>%
  filter(Site.code != "OLCAP5B")

# loading jacknife scores 
jackAll <- read.csv("results/jackAll.csv")
jack_rainForest <- read.csv("results/jackRainforest.csv")
jack_openCountry <- read.csv("results/jackOpencountry.csv")

# Load vegetation data from previous scripts 
vegData <-  read.csv("results/summaryVeg.csv") %>%
  filter(!str_detect(Site_ID, 'OLCAP5B'))
vegPcaScores <- read.csv("results/pcaVeg.csv") %>%
  filter(!str_detect(Site_ID, 'OLCAP5B'))

# proportion of acoustic detections of rainforest and open-country birds across visits
propVisit <- read.csv("results/acoustic-detections-across-visits.csv")
# convert date column to character for left_join
propVisit$Date <- as.character(propVisit$Date)

# load the entire asu data across all sites and days computed
sitebyDayAsu <- read.csv("results/site-by-day-asu.csv") 

# separate by Site and Date
sitebyDayAsu <- separate(sitebyDayAsu, col = Site_Day, into = c("Site", "Date"), sep = "_") 

# Add restoration type column to the space use data
sitebyDayAsu <- left_join(sitebyDayAsu, sites, by=c("Site"="Site.code"))

# scale values per site/date for comparison between sites and treatment types
sitebyDayAsu <- sitebyDayAsu %>%
  group_by(Site, Date, Restoration.type) %>%
  mutate(f.cont.scaled = range01(f.cont))

# Let's look at data by restoration type
# This suggests that we have more data for benchmark sites relative to the other two treatment types
nDays_siteType <- sitebyDayAsu %>%
  dplyr::select(Site, Date, Restoration.type) %>%
  distinct() %>%
  group_by(Restoration.type) %>%
  count()
  
# Prepare data for statistical modeling
# Calculating total space use across all frequency bins and times of day: 128*24 for each site-day combination

totSpaceUse <- sitebyDayAsu %>%
  group_by(Site, Date, Restoration.type) %>%
  summarise(totSpaceuse = sum(f.cont.scaled)) %>%
  group_by(Site) %>% 
  mutate(visit = row_number()) %>%
  mutate(siteCode = str_extract(Site, pattern = "\\w+\\d+")) %>%
  mutate(siteCode = factor(siteCode)) %>%
  full_join(vegData, by=c("Site"="Site_ID")) %>%
  ungroup()
```


### Getting data ready in a format for linear modeling
```{r}
# overall space use and bird species detections
modelDataAll <- vegPcaScores %>% 
  rename(Site = Site_ID) %>%
  rename(Restoration.type = Site_type) %>%
  mutate(across(Restoration.type, factor))  %>%
  full_join(totSpaceUse, by=c("Site","Restoration.type")) %>%
  mutate("roundSpaceuse" = round(totSpaceuse)) %>%
  full_join(propVisit[,-2], by=c("Site","Restoration.type")) %>%
  distinct(.)
```

### Acoustic space use and proportion of bird species detections
```{r}
# rainforest bird species detections
glmm_detectionsRF_space <- glmer(roundSpaceuse ~ propRF + (1|siteCode) + (1|visit), data = modelDataAll, family = gaussian(link="identity"))

summary(glmm_detectionsRF_space)
plot_model(glmm_detectionsRF_space, type="pred")
report::report(glmm_detectionsRF_space)

# We fitted a linear mixed model (estimated using REML and nloptwrap optimizer) to predict roundSpaceuse with propRF (formula: roundSpaceuse ~ propRF). The model included siteCode and visit as random effects (formula: list(~1 | siteCode, ~1 | visit)). The model's total explanatory power is substantial (conditional R2 = 0.84) and the part related to the fixed effects alone (marginal R2) is of 1.99e-04. The model's intercept, corresponding to propRF = 0, is at 289.91 (95% CI [233.66, 346.16], t(1215) = 10.11, p < .001). Within this model:

#  - The effect of propRF is statistically non-significant and positive (beta = 15.96, 95% CI [-18.66, 50.57], t(1215) = 0.90, p = 0.366; Std. beta = 0.01, 95% CI [-0.02, 0.05])

# Standardized parameters were obtained by fitting the model on a standardized version of the dataset. 95% Confidence Intervals (CIs) and p-values were computed using the Wald approximation.

# open-country bird species detections
glmm_detectionsOC_space <- glmer(roundSpaceuse ~ propOC + (1|siteCode) + (1|visit), data = modelDataAll, family = gaussian(link="identity"))

summary(glmm_detectionsOC_space)
plot_model(glmm_detectionsOC_space, type="pred")
report::report(glmm_detectionsOC_space)

# We fitted a linear mixed model (estimated using REML and nloptwrap optimizer) to predict roundSpaceuse with propOC (formula: roundSpaceuse ~ propOC). The model included siteCode and visit as random effects (formula: list(~1 | siteCode, ~1 | visit)). The model's total explanatory power is substantial (conditional R2 = 0.84) and the part related to the fixed effects alone (marginal R2) is of 1.99e-04. The model's intercept, corresponding to propOC = 0, is at 305.87 (95% CI [257.26, 354.48], t(1215) = 12.35, p < .001). Within this model:

#  - The effect of propOC is statistically non-significant and negative (beta = -15.96, 95% CI [-50.57, 18.66], t(1215) = -0.90, p = 0.366; Std. beta = -0.01, 95% CI [-0.05, 0.02])

# Standardized parameters were obtained by fitting the model on a standardized version of the dataset. 95% Confidence Intervals (CIs) and p-values were computed using the Wald approximation. 
```

### Year since restoration

Let's look at year since restoration and its effect on acoustic space use and the proportion of rainforest and open country bird species detections
```{r}
# prep dataframe
allRestored <- modelDataAll %>%
  filter(Restoration.type=="Active") %>%
  mutate(yearSinceRestoration = (2022-plantingYear))

## acoustic space use
glmmYearSpace <- glmer(totSpaceuse ~ yearSinceRestoration + (1|visit), data = allRestored, family = gaussian(link="identity"))
summary(glmmYearSpace)

plot_model(glmmYearSpace, type="pred")
report::report(glmmYearSpace)

# We fitted a linear mixed model (estimated using REML and nloptwrap optimizer) to predict totSpaceuse with yearSinceRestoration (formula: totSpaceuse ~ yearSinceRestoration). The model included visit as random effect (formula: ~1 | visit). The model's total explanatory power is moderate (conditional R2 = 0.16) and the part related to the fixed effects alone (marginal R2) is of 0.12. The model's intercept, corresponding to yearSinceRestoration = 0, is at 343.91 (95% CI [238.78, 449.04], t(61) = 6.54, p < .001). Within this model:

#  - The effect of yearSinceRestoration is statistically significant and negative (beta = -10.07, 95% CI [-16.66, -3.47], t(61) = -3.05, p = 0.003; Std. beta = -0.35, 95% CI [-0.58, -0.12])

# Standardized parameters were obtained by fitting the model on a standardized version of the dataset. 95% Confidence Intervals (CIs) and p-values were computed using the Wald approximation.

## rainforest bird species detections
glmmYearRF <- glmer(propRF ~ yearSinceRestoration + (1|visit), data = allRestored, family = gaussian(link="identity"))
summary(glmmYearRF)

plot_model(glmmYearRF, type="pred")
report::report(glmmYearRF)

# We fitted a linear mixed model (estimated using REML and nloptwrap optimizer) to predict propRF with yearSinceRestoration (formula: propRF ~ yearSinceRestoration). The model included visit as random effect (formula: ~1 | visit). The model's explanatory power related to the fixed effects alone (marginal R2) is 3.67e-09. The model's intercept, corresponding to yearSinceRestoration = 0, is at 0.77 (95% CI [0.71, 0.83], t(366) = 24.62, p < .001). Within this model:

#  - The effect of yearSinceRestoration is statistically non-significant and negative (beta = -2.33e-06, 95% CI [-3.94e-03, 3.93e-03], t(366) = -1.16e-03, p > .999; Std. beta = -6.07e-05, 95% CI [-0.10, 0.10])

# Standardized parameters were obtained by fitting the model on a standardized version of the dataset. 95% Confidence Intervals (CIs) and p-values were computed using the Wald approximation. 

## open country bird species detections
glmmYearOC <- glmer(propOC ~ yearSinceRestoration + (1|visit), data = allRestored, family = gaussian(link="identity"))
summary(glmmYearOC)

plot_model(glmmYearOC, type="pred")
report::report(glmmYearOC)

# We fitted a linear mixed model (estimated using REML and nloptwrap optimizer) to predict propOC with yearSinceRestoration (formula: propOC ~ yearSinceRestoration). The model included visit as random effect (formula: ~1 | visit). The model's explanatory power related to the fixed effects alone (marginal R2) is 3.67e-09. The model's intercept, corresponding to yearSinceRestoration = 0, is at 0.23 (95% CI [0.17, 0.29], t(366) = 7.32, p < .001). Within this model:

#  - The effect of yearSinceRestoration is statistically non-significant and positive (beta = 2.33e-06, 95% CI [-3.93e-03, 3.94e-03], t(366) = 1.16e-03, p > .999; Std. beta = 6.07e-05, 95% CI [-0.10, 0.10])

# Standardized parameters were obtained by fitting the model on a standardized version of the dataset. 95% Confidence Intervals (CIs) and p-values were computed using the Wald approximation. 
```
